---
title: "Simulation Exercise"
author: "GS"
date: "4/8/2021"
output: pdf_document
---

```{r setup, include=FALSE, comment='', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.


```{r, message=FALSE}
library(tidyverse)
library(tinytex)
theme_set(theme_bw())
```

## Simulations
Setting lambda y exponentials
```{r}
set.seed(2021) #for reproducibility

lambda = .2
expo = 40

sim_means = NULL
for (i in 1 : 1000) sim_means = c(sim_means, mean(rexp(expo, lambda)))
```

## Theoretical mean vs Sample mean
Sample mean (from the simulation)
```{r}
print(paste('The sample mean is ', round(mean(sim_means), 4)))

print(paste('The theoretical mean is ', 1/lambda))

mean_dif <- abs(mean(sim_means) - (1/lambda))
print(paste('The mean difference is ', round(mean_dif, 4)))
```
## Theoretical variance vs Sample variance
```{r}
print(paste('The sample variance is ', round(var(sim_means), 4)))

print(paste('The theoretical variance is ', (lambda * sqrt(expo))^-2))

var_dif <- var(sim_means) - (lambda * sqrt(expo))^-2
print(paste('The variance difference is ', round(var_dif, 4)))
```
```{r}
ggplot(data.frame(y=sim_means), aes(x=y)) +
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="lightblue",
                 color="black") +
  stat_function(fun=dnorm, args=list(mean=1/lambda,
                                    sd=(lambda*sqrt(expo))^-1), size=2) +
  xlab('Mean') + 
  ggtitle('Simulations')
```

```{r}
sessionInfo()
```

